Controlling the dewetting morphologies of thin liquid films by switchable substrates

### 2023-07-10 Reply_RefereeC.tex

\section*{Reply to Referee C}
We thank the Referee for his/her report on our manuscript.
We are glad to read that he/she finds the study scientifically sound and credible and 
the manuscript well written. Based on his/her important remarks, we have 
prepared a revised version which we hope will be convincing also for what concerns
its impact and interest for the readership of Physical Review Letters.
Hereafter, detailed answers to the various questions raised are reported. All changes 
in the text are edited in red color, for the sake of clarity.

\begin{itemize}

\item[ \textbf{\underline{Comment 1.}}]
{ 
However, I feel that the
results presented lack a sufficient breadth and impact so as to
warrant publication in Physical Review Letters - provided that one has
a suitable substrate, I do not see how would the transport of a fluid
by way of transient rivulets surpass the transport by way of droplets
in practice.
}

\item[ \textbf{{Answer}}]
{ 
It is not stated in the paper that the rivulets would transport fluid more efficiently and actually we do not raise any statements about transport at all. The focus of our contribution is not on sliding or similar force-driven processes, but 
on dewetting and, in particular, on the different dewetting {\it morphologies}. 
It is correct, on the other hand, that rivulets are metastable. However, as 
we showed, their lifetime can be extended by increasing the pattern speed
(therefore enabling treatment, e.g. curing, of the substrate to 
fix a given pattern).
We think, though, that probably the origin of this misunderstanding could be in the use of 
the word 'transport' when we introduce the parameter $\Gamma$: we wrote, in fact, 
"{\it if $U_{\Theta}$ is large as compared to $v_{\theta}$, the film retraction is faster than 
the transport due to the contact angle field and thus droplets from}". We agree that such use 
in that context might be not appropriate and misleading and we apologize for it. We have, thus,
rephrased the sentence as follows:\\
\\
\textcolor{red}{If $U_{\Theta}$ is large as compared to $v_{\theta}$, the film retraction is faster than 
the local contact angle variation and thus droplets form.}\\
\\
On the other hand, if the Referee does not believe that it is possible to actually achieve a fast 
contact angle variation, as compared to a typical retraction speed, i.e. a large $\Gamma$, we wrote a new paragraph, just before the conclusions, where we envisage and discuss a possible experimental realization of our numerical setup, that would allow to easily get $\Gamma \approx 10^3$.\\
The breadth and impact to warrant publication in Physical Review Letters is in our opinion clearly given: As we stressed extensively in the introduction,
the problem of controlling dewetting is crucially relevant for applications in 
nanoprinting (including printable electronics) or the development and application of new smart materials. Here, we provide a new approach to control the dewetting morphology by means of switchable substrates which might enable new deposition techniques allowing for unprecedented control. 
Besides the possible application impact, the problem is of fundamental interest in Fluid Dynamics (which represents an important part of the 
PRL readership).
This is witnessed by the various publications in PRL on this and closely related problems. To cite but a few recent ones: J.C. Fern\'andez-Toledano et al, "How wettability controls nanoprinting",
{\it Phys. Rev. Lett.} {\bf 124} 224503 (2020); A.A. Pahlavan et al,
"Evaporation of binary-mixture liquid droplets: the formation of picoliter pancakelike shapes", {\it Phys. Rev. Lett.} {\bf 127}, 024501 (2021); C. Clavaud et al, "Modification of the fluctuation dynamics of ultrathin wetting films", {\it Phys. Rev. Lett.} {\bf 126}, 228004 (2021).
}

### 2023-07-10 Reply_RefereeD.tex

\documentclass[12pt,english]{article}

\usepackage{natbib}

\usepackage{graphics,graphicx,dcolumn,bm,fleqn,epic,eepic,float}
\usepackage{amssymb,amsmath,multirow,rotate,rotating,color}
\usepackage[utf8]{inputenc}

\usepackage[english]{babel}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tikz}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}
%\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\tikzset{fontscale/.style = {font=\relsize{#1}}}
\usetikzlibrary{calc}
\makeatother

\newcommand{\figref}[1]{Fig.~\ref{fig:#1}}
\newcommand{\eqnref}[1]{Eq.~(\ref{eq:#1})}
\newcommand{\ts}{\textsuperscript}

\definecolor{tuered}{RGB}{214,0,74}

\newcommand{\todo}[1]{\textbf{\textcolor{tuered}{ TODO: #1}}}

\newcommand{\vectornorm}[1]{\left|\left|#1\right|\right|}

\renewcommand{\vec}[1]{\mathbf{#1}}
\newcommand{\uvec}[1]{\hat{\vec{#1}}}
\newcommand{\tensor}[1]{\mathbf{#1}}

\definecolor{pyblue}{HTML}{1F77B4}
\definecolor{pyorange}{HTML}{FF7F0C}
\definecolor{pygreen}{HTML}{2CA02C}
\definecolor{pyred}{HTML}{D62728}

\newcommand{\JH}[1]{\textcolor{blue}{JH: #1}}
\begin{document}

% The referee PDF has been generated from r239 -- 'make referee' command.

\section*{Reply to Referee D}
We thank the new Referee for his/her report on our manuscript. 
We are glad to read that he/she acknowledges the systematic approach we have taken and are happy to see that he/she finds our results interesting. 
We understand the concerns raised by the referee and wonder why the recommendation is negative.
Hereafter, we provide answers to the various questions/comments raised. 
All changes in the revised version are reported in red color.

\begin{itemize}

\item[ \textbf{\underline{Comment 1.}}]
{ 
It seems impossible to construct the patterns described by Equation (3) in a real experiment, making the current research less attractive.
}

\item[ \textbf{{Answer}}]
{
We are very surprised by this comment and do not agree with the assessment of the referee. 
% For static patterns it is well known how to pattern substrates in a precise and controlled way using, e.g. lithography. 
It is true that the novelty of the here presented work is strongly coupled to both the temporal and spatial component of the wettability and as such to equation (3).
We are certain that switchable substrates offer this kind of control, and for photoswitchable materials spiropyran should be a perfect candidate~\cite{keyvanradSpiropyranbasedAdvancedPhotoswitchable2022}.

Not only do we present our numerical and analytical results in this letter, but we also supply a potential realization for a lab experiment. 
We provide a schematic setup using a photoswitchable substrate and a digital multimirror device. 
As we are well aware that this is not sufficient for an experiment we further identify the relevant length and time scales for both the film and the apparatus.
% Hoping that experimentalist confirm transition we supply a way to construct pattern Eq. (3).
%We therefore have a hard time understanding why it seems impossible to the referee to construct the pattern equation %(3), as he/she does not give us a single increment of information in his/her comment.
}

\item[ \textbf{\underline{Comment 2.}}] 
{
This work is completed based on the lubrication equation for the thin film flows.
From theoretical/mathematical aspects, the novelty of this work, in my opinion, is not strong enough. 
This lubrication model employed here was proposed about 40 years ago and has been thoroughly investigated~\cite{RevModPhys.69.931, RevModPhys.81.1131}. 
There are no improvements or modifications to the model in this work. 
Additionally, the model for the rupture time comes from a simple scaling analysis, which is not very convincing when considering other classical mathematical approaches, such as linear instability analysis, similarity transformation, etc.
}

\item[ \textbf{{Answer}}]
{
We find it puzzling that the referee arrives at the conclusion that working on a theory which is 40 years old does not allow for the novelty attributed to PRL.
A good counter example is: M. Wilczek et al,
"Sliding drops – ensemble statistics from single drop bifurcations", {\it Phys. Rev. Lett.} {\bf 119}, 204501 (2017), in which Wilczek et al. use the thin film equation on an inclined plane.
Similar arguments can be constructed about the Navier-Stokes equation which is 180 years old, still there is a lot of research dedicated towards, e.g., turbulence.

We agree with the referee that from a purely mathematical point of view lubrication theory is untouched.
There is no reason or advantage to add artificial complexity to a well working theory which has been developed for this singular purpose.
However, when it comes to dynamical effects and non-linear behaviour of thin films, the topic this letter revolves around, mathematics falls short and only numerical simulation or experiments offer deeper insights.
To this end we use a novel lattice Boltzmann method to solve the thin film equation at time scales where initial small perturbations have grown to similar sizes as $h_0$.
It is well known that the rupture time marks the end of the linear regime and linear instability theory as suggested by the referee, which has been used in our previous work~\cite{PhysRevE.104.034801}. 
However, it is simply not able to explain our observations, as $\tau_r$ is dependent on $\Gamma$. 
Observing the morphological transition during dewetting from a droplet to a rivulet state is by no means covered by linear analysis. 
It is a dynamic competition of relevant velocities which to the best knowledge of the authors has not been discussed or shown before.
}

\item[ \textbf{\underline{Comment 3.}}]
{
The mobility function in Equation (1) contains the slip length $\delta$. 
But there is no further information in this work. 
Is the slip length of a patterned substrate constant? 
If so, it is quite unreasonable because a hydrophobic substrate is usually more slippery than a hydrophilic one. 
Moreover, previous works had showed that the slip is crucial to the interface dynamics~\cite{zhang_sprittles_lockerby_2021, zhao_zhang_si_2023}, which should be taken into account.
}

\item[ \textbf{{Answer}}]
{
We thank the Referee for this remark. It is of course true that the slip length 
should grow with the contact angle, on the other hand it is also well known that for low contact angle (actually up to $\sim 100^{\circ}$) such dependence is relatively weak, 
as it is expected to follow the relation $\lambda \propto (1+\cos \theta)^{-2}$
(see Huang et al, {\it Phys. Rev. Lett.} {\bf 101}, 226101 (2008)). The ratio 
of the slip lengths on patches of contanct angles $10^{\circ}$ and $30^{\circ}$
(which is the case of our simulations), for instance, would then be $\approx 0.9$,
such that the slip length should have been varied between $0.95$ and $1.05$ lbu. We deemed, then, reasonable to consider it approximately constant. 
To provide an idea, in reference~\cite{zhang_sprittles_lockerby_2021}, for instance, effects were appreciated 
when slip lengths were two to five times larger than each other. 
}

\item[ \textbf{\underline{Comment 4.}}]
{ 
According to the fluid parameters displayed in the supplemental material, the initial thickness of the film is about 4 nm. 
Thermal fluctuations had been shown to play a key role at this scale~\cite{PhysRevLett.99.114503, PhysRevLett.126.228004}, which might affect the dynamics seriously.
}

\item[ \textbf{{Answer}}]
{
We do not say that the initial film thickness is $4 \, \text{nm}$, we just considered the experiments of Becker et al (where the initial film thickness is $\sim 4 \, \text{nm}$) as a reference, since they provide experimental values of quantities (such as the Hamaker constant) that enable the comparison 
with the numerics. The effect of thermal fluctuations, though, is known to essentially show up as 
a speed up of the growth of the instability, but it does not affect particularly the dewetting morphology (as shown in Becker et al, {\it Nat. Mater.} {\bf 2}, 59--63, who made direct comparison 
of the numerical simulation of a lubrication model with experiments).
}

\end{itemize}

\bibliographystyle{abbrv}
\bibliography{Ref}

\end{document}

### 2023-07-10 main_exciting.tex
More excitement

\begin{abstract}
\noindent Switchable and adaptive substrates emerged as valuable tools for controlling wetting and actuation of droplet motion. Here, we report a computational study of the dynamics of an unstable thin liquid film deposited on a switchable substrate, modeled with a space and time varying contact angle.
With a static pattern, all fluid is drained into droplets located around contact angle minima, whereas for a sufficiently large rate of wettability variation a state consisting of metastable rivulets is observed. 
A criterion discriminating whether rivulets occur or not is identified in terms of a single dimensionless parameter.
Finally, we show and explain theoretically how the film rupture times, droplet shape and rivulet life time depend on the pattern wavelength and speed.  
\end{abstract}

>>>>

\begin{abstract}
\noindent Switchable and adaptive substrates emerged as valuable tools for controlling wetting and actuation of droplet motion. Here, we report a computational study of the dynamics of an unstable thin liquid film deposited on a switchable substrate, modeled with a space and time varying contact angle.
With a static pattern, all fluid is drained into droplets located around contact angle minima, whereas for a sufficiently large rate of wettability variation a \REV{novel topological transition as such the occurence of a metastable rivulet state is observed}. 
A criterion discriminating whether rivulets occur or not is identified in terms of a single dimensionless parameter.
Finally, we show and explain theoretically how the film rupture times, droplet shape and rivulet life time depend on the pattern wavelength and speed.  
\end{abstract}
